Realizing temporal graphs from fastest travel times

### 2023-09-22 STACS2023/STACS23-shortversion.tex

Unfortunately the above guesses do not provide us with enough information to be able to determine all fastest paths. For example consider the case depicted in the FIGURE TODO. 
Therefore we introduce some more guesses. To do this we have to first define the following.
\begin{definition}\label{def:segments}
	Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
	A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for all $i \in \{ 2, 3, \dots, p-1\}$,
	is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
\end{definition}
Note by \cref{def:segments} that $S_{u,v} \neq S_{v,u}$. 
Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or it consists of a sequence of some segments. Furthermore, since we have at most $4k$ interesting vertices in $G'$, we can deduce the following important result.
\begin{corollary}\label{obs:FPT-k2segments}
	There are at most $O(k^2)$ segments in $G'$.
\end{corollary}

@@

There are two more guesses \textcolor{lipicsGray}{\textsf{\textbf{G-11}}} and \textcolor{lipicsGray}{\textsf{\textbf{G-12}}} that are deferred to the Appendix. We prove that there are $f(k)$ combinations of guesses and for each one we create an ILP with $f(k)$ variables and $f(k)\cdot |D|^{O(1)}$ constraints.
%We create all of the above guesses independently.
%We start with the first one, that results in $O(k^{k^3})$ different possibilities, then we split each one of these guesses into $O(k^{k^3})$ new ones, that respond to the guessing in the second step, etc.
%After creating all of the guesses we end up with $f(k)$ different cases (where $f$ is a double exponential) 
%and create one ILP instance for each case.
%
%For each of the ILP instances we iterate through all pairs of vertices $u,v \in V(G)$ and, using the above guesses, determine, not only the fastest path $P$ from $u$ to $v$, but also all other paths $P'_i$.
%For the labels of edges of the fastest path $P$ we introduce the equality constraint, while for the labels of the edges of all other paths $P'_i$ we introduce the inequality constraint (\ie the duration $d(P_i')$ of any $P'_i$ cannot be smaller than $d(P)=D_{u,v}$).
%We can do this efficiently, while introducing only a small number of variables,
%which ensures that each ILP is solved in an FPT-time.
For detailed explanation and proofs of this part see Appendix.

>>>>

\begin{figure}[t]
	\centering
	%\includegraphics{fig-exampleg10}
	\caption{In the above graph vertices $v_1, v_{11}, w$ are in $U$, while $v_2, v_{10}$ are in $U^*$. 
		Numbers above all $v_i$ represent the values of the fastest temporal paths from $w$ to each of them (\ie the entries in the $w$-th row of matrix $D$).
		From the basic guesses we know the fastest temporal path $P$ from $w$ to $v_2$ (depicted in blue) and the fastest temporal path $Q$ from $w$ to $v_{10}$.
		From the values of durations from $w$ to each $v_i$ we cannot 
		determine the fastest paths from $w$ to all $v_i$.
		More precisely, we know that $w$ reaches $v_2, v_3, v_4, v_5$ (resp.~$v_{10}, v_{9}, v_{9}, v_{7}$) 
		by first using the path $P$ (resp.~$Q$) and then proceeding through the vertices,
		but we do not know how $w$ reaches $v_6$ the fastest.
		Therefore we have to introduce some more guesses.
  \label{fig:guesses-advanced}}
\end{figure}

Unfortunately the above guesses do not provide us with enough information to be able to determine all fastest paths. For example consider the case depicted in \cref{fig:guesses-advanced}. 
Therefore we introduce some more guesses. To do this we have to first define the following.
\begin{definition}\label{def:segments}
	Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
	A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for all $i \in \{ 2, 3, \dots, p-1\}$,
	is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
\end{definition}
Note by \cref{def:segments} that $S_{u,v} \neq S_{v,u}$. 
Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or it consists of a sequence of some segments. Furthermore, since we have at most $4k$ interesting vertices in $G'$, we can deduce the following important result.
\begin{corollary}\label{obs:FPT-k2segments}
	There are at most $O(k^2)$ segments in $G'$.
\end{corollary}

@@

There are two more guesses \textcolor{lipicsGray}{\textsf{\textbf{G-11}}} and \textcolor{lipicsGray}{\textsf{\textbf{G-12}}} that are deferred to the Appendix. 
We prove in the Appendix that, for all guesses \textcolor{lipicsGray}{\textsf{\textbf{G-1}}} to \textcolor{lipicsGray}{\textsf{\textbf{G-12}}}, there are in total at most $f(k)$ possible choices, and for each one of them we create an ILP with at most $f(k)$ variables and at most $f(k)\cdot |D|^{O(1)}$ constraints. Each of these ILPs can be solved in FPT time by Lenstra's Theorem~\cite{Lenstra1983Integer}.
%We create all of the above guesses independently.
%We start with the first one, that results in $O(k^{k^3})$ different possibilities, then we split each one of these guesses into $O(k^{k^3})$ new ones, that respond to the guessing in the second step, etc.
%After creating all of the guesses we end up with $f(k)$ different cases (where $f$ is a double exponential) 
%and create one ILP instance for each case.
%
%For each of the ILP instances we iterate through all pairs of vertices $u,v \in V(G)$ and, using the above guesses, determine, not only the fastest path $P$ from $u$ to $v$, but also all other paths $P'_i$.
%For the labels of edges of the fastest path $P$ we introduce the equality constraint, while for the labels of the edges of all other paths $P'_i$ we introduce the inequality constraint (\ie the duration $d(P_i')$ of any $P'_i$ cannot be smaller than $d(P)=D_{u,v}$).
%We can do this efficiently, while introducing only a small number of variables,
%which ensures that each ILP is solved in an FPT-time.
For detailed explanation and proofs of this part see Appendix.
